# Quick-start guide

## Set up environment
### Load flycircuit package
```{r load-flycircuit, cache=T}
library(flycircuit)
```


### Load visual projection neuron NBLAST scores
Download NBLAST scores for visual projection neurons and load them into the workspace.

**WARNING:** The download totals 146 MB.
```{r download-vpn_scores, cache=T}
# Download scores...
fc_download_data('http://flybrain.mrc-lmb.cam.ac.uk/flycircuit/vpn_scores.desc', type='bigmat')
```
```{r load-vpn_scores, cache=T}
# ...and load them into the workspace
fc_attach_bigmat('vpn_scores')
```

Show the scores for the first five genes:
```{r score-peek}
vpn_scores[1:5, 1:5]
```


### Download visual projection neuron data
At the moment we have the NBLAST scores for the visual projection neurons, but no other information about them. Let's download some.
```{r download-vpn-info, cache=T}
# Download the data and save it to a 'vpns' folder in the home directory.
vpns <- read.neuronlistfh('http://flybrain.mrc-lmb.cam.ac.uk/flycircuit/vpns.rds', '~/vpns')
```
This `vpns` object now contains summary data about all `r length(vpns)` the visual projection neurons. When more information about a particular neuron is requested, the corresponding file is downloaded and saved to a `data` subdirectory of the folder specified in the `read.neuronlistfh` command.


### Plotting NBLAST hits
Now, let's plot the two neurons that NBLAST thinks are most similar.
```{r plot top-hits, cache=T}
# Find the indices of the maximum score in the score matrix
ind <- which(vpn_scores[, ] == max(vpn_scores[, ]), arr.ind=T)
queryneuron <- colnames(vpn_scores)[ind[2]]
targetneuron <- rownames(vpn_scores)[ind[1]]

# What was the query?
queryneuron

# What was the target?
targetneuron

# What do they look like?
open3d()
plot3dfc(c(queryneuron, targetneuron), db=vpns)

# Where in the brain are they?
fcwbsurf()
```
